---
title: "Debug logging example"
output: md_document
---

**Objective:** Enable printf-style debugging of recursive function calls with minimal code.

Let's define a simple recursive debug logger.
The logger will print function name, argument names and values, and return value.
We want the logger to be minmally invasive so that it is easy to annotate a function as debugged.
```{r}
.GlobalEnv$.log.indent <- 0
`%logged%` <- function(prefix, fun) {
    function (...) {
        prefix <- paste0(prefix, paste0(rep('  ', .GlobalEnv$.log.indent), collapse=''))
        cat(prefix, deparse(match.call()), '\n')
        cat(prefix, '>', capture.output(print(list(...))), '\n')
        .GlobalEnv$.log.indent <- .GlobalEnv$.log.indent + 1
        ret <- fun(...)
        .GlobalEnv$.log.indent <- .GlobalEnv$.log.indent - 1
        cat(prefix, '<', capture.output(print(ret)), '\n')
        return(ret)
    }
}
```

To use our debugger, we need only prepend `'DEBUG' %logged%` to each function we want to debug. Suppose we start with a function
```{r}
fib <- function (x) {
    if (x == 0 || x == 1) return(x)
    return(fib(x - 1) + fib(x - 2))
}
fib(5)
```
To debug `fib`, we'll simply add `'DEBUG' %logged%` to its definition (typically deep in some .R file)
```{r}
fib <- 'DEBUG' %logged% function (x) {  # <--- only this line changed
    if (x == 0 || x == 1) return(x)
    return(fib(x - 1) + fib(x - 2))
}
fib(5)
```
Here the `>` lines print the input arguments to `fib()` and the `>` print its return value. Since `fib()` calls itself, we also see intermediate logging between `>` and `<`.

If you want to log two different subsystems, you can give them different prefixes instead of just `'DEBUG'`, but all prefixes should have the same number of characters.