---
title: "Trace logging example"
output: md_document
---

**Objective:** Log an selective execution trace to a file, so that traces of two different executions can be diffed.

# Defining a tracing tool

Let's define a simple recursive tracing tool.
```{r}
.GlobalEnv$.trace.file = ''
.GlobalEnv$.trace.filename = NA
tracer.start <- function(filename = '/tmp/trace.txt') {
    .GlobalEnv$.trace.filename <- filename
    .GlobalEnv$.trace.file <- file(filename, open = 'w+')
}
tracer.stop <- function() {
    close(.GlobalEnv$.trace.file, type = 'w+')
    .GlobalEnv$.trace.file <- file(.GlobalEnv$.trace.filename, open = 'r')
    lines <- readLines(.GlobalEnv$.trace.file)
    close(.GlobalEnv$.trace.file, type = 'r')
    .GlobalEnv$.trace.file = ''
    return(lines)
}
tracer.enter <- function() {
    call <- deparse(sys.call(sys.parent(1)))
    if (length(call) > 1) {
        call <- paste(call[[1L]], "....")
    }
    cat(paste(rep(' ', .GlobalEnv$.log.indent)), call, "\n", file = .GlobalEnv$.trace.file)
    .GlobalEnv$.log.indent <- .GlobalEnv$.log.indent + 1
}
tracer.exit <- function() {
    .GlobalEnv$.log.indent <- .GlobalEnv$.log.indent - 1
}
tracer.trace <- function(...) {
    .GlobalEnv$.log.indent <- max(0, .GlobalEnv$.log.indent)
    untrace(...)
    trace(..., tracer.enter, tracer.exit, print = FALSE)
}
```

To use our debugger, we need only prepend `'DEBUG' %logged%` to each function we want to debug. Suppose we start with a function
```{r}
fib <- function (x) {
    if (x == 0 || x == 1) return(x)
    return(fib(x - 1) + fib(x - 2))
}
fib(5)
```
To debug `fib`, we'll simply add `'DEBUG' %logged%` to its definition (typically deep in some .R file)
```{r}
tracer.trace(fib)
fib(5)
```
Next let's write `fib`'s trace to file.
```{r}
tracer.start('/tmp/trace.fib.txt')
tracer.trace(fib)
fib(5)
lines <- tracer.stop()
```
Now we can print the trace
```{r}
cat(lines, sep = '\n')
```

# Tracing the NIMBLE compilation of two different functions.

**Objective:** I'm hunting down a bug in an initial prototype of `optim()` in NIMBLE.

In this section we'll diff two different `compileNimble` invocations: a successful run using `eigen` and a failing run using `optim`.
```{r}
library(nimble)
testCompilation <- function(type) {
    # Setup a nimbleFunction to compile.
    if (type == 'optim') {
        run <- function(par = double(1)) {
            return(fakeOptim(par))
            returnType(optimResultNimbleList())
        }
    } else if (type == 'eigen') {
        run <- function(par = double(2)) {
            return(eigen(par))
            returnType(eigenNimbleList())
        }
    } else {
        stop(paste('Unknown type:', type))
    }
    nimFun <- nimbleFunction(setup = TRUE, run = run)()
    filename <- paste0('/tmp/trace.', type, '.txt')

    # Configure which internal NIMBLE functions to trace.
    # TODO add more functions to trace here, so that the two traces differ.
    tracer.trace(compileNimble)
    # tracer.trace(nimble:::nimbleProjectClass)  # FIXME this does not work.

    # Run the function, dumping trace to file.
    tracer.start(filename)
    try(compileNimble(nimFun, showCompilerOutput = TRUE, dirName = '~/tmp'))
    tracer.stop()
    return(filename)
}
```
Now we'll run both versions.
```{r}
eigenFile <- testCompilation('eigen')
optimFile <- testCompilation('optim')
```
Finally, we'll open the two traces in an external diff tool
```{r}
system2('meld', c(eigenFile, optimFile), wait = FALSE)
```

